---
title: "scRNA w/ Seurat"
output: html_document
date: "2024-02-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library, include=FALSE}
library(dplyr)
library(Seurat)
library(patchwork)
library(sctransform)
library(DoubletFinder)
```

## Loading Data

10k Human PBMC, 3' Human peripheral blood mononuclear cells of a healthy female donor aged 25-30.

Libraries were generated from ~16,000 cells (11,984 recovered). Sequenced on an illumina novaseq 6000 to read depth of approximately 40,000 mean reads per cell. 


```{r data_loading}
# Load the PBMC dataset
pbmc.data <- Read10X(data.dir = "/home/me/rotation_ii/week2/filtered_feature_bc_matrix")
# Initialize the Seurat object with the raw (non-normalized data).
pbmc <- CreateSeuratObject(counts = pbmc.data, min.cells = 3, min.features = 200)
# store mitochondrial percentage in object meta data
pbmc <- PercentageFeatureSet(pbmc, pattern = "^MT-", col.name = "percent.mt")
```
# Plan: Compare four different pipelines:

For this markdown file we are going to run four pipelines:

I) Seurat Standard
II) Seurat Standard w/ Doublet Finder
III) SCTransform
IV) SCTransform w/ Doublet Finder

# Seurat Standard Pipeline
### Visualizing QC metrics 
### Filtering based on QC metrics
### Normalization
### Feature selection
### Scaling data
### PCA
### KNN
### Clusters
### UMAP/tSNE

```{r}
# Visualize QC metrics as a violin plot
VlnPlot(pbmc, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```

```{r}
plot1 <- FeatureScatter(pbmc, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(pbmc, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1 + plot2
```

```{r}
pbmc <- subset(pbmc, subset = nFeature_RNA > 200 & nFeature_RNA < 80000 & percent.mt < 25)
pbmc <- NormalizeData(pbmc)
pbmc <- FindVariableFeatures(pbmc, selection.method = "vst", nfeatures = 2000)
pbmc <- ScaleData(pbmc)
pbmc <- RunPCA(pbmc)
```


```{r}
ElbowPlot(pbmc)
```

I would choose a PC of 10 here. 
```{r}
PC_count = 10
pbmc <- FindNeighbors(pbmc,dims=1:PC_count)
pbmc <- FindClusters(pbmc,resolution=1.2)
```
```{r}
pbmc <- RunUMAP(pbmc,dims=1:10)
DimPlot(pbmc,reduction="umap")
```

```{r}
saveRDS(pbmc, file = "/home/me/rotation_ii/week2/Seurat_nodoublet.rds")
```

```{r}
pbmc <- readRDS( "/home/me/rotation_ii/week2/Seurat_nodoublet.rds")
```







# Seurat Standard Normalization with DoubletFinder
### pK Identification and Doublet Finder with Standard Seurat Pipeline

```{r}
#sweep.res.list_blood <- paramSweep(pbmc)
#sweep.stats_blood <- summarizeSweep(sweep.res.list_blood)
#bcmvn_blood <- find.pK(sweep.stats_blood)
```





# SCTransform Normalization


```{r data_loading}
# Load the PBMC dataset
pbmc.data <- Read10X(data.dir = "/home/me/rotation_ii/week2/filtered_feature_bc_matrix")
# Initialize the Seurat object with the raw (non-normalized data).
pbmc <- CreateSeuratObject(counts = pbmc.data, min.cells = 3, min.features = 200)
# store mitochondrial percentage in object meta data
pbmc <- PercentageFeatureSet(pbmc, pattern = "^MT-", col.name = "percent.mt")
```

```{r}
pbmc <- SCTransform(pbmc, vars.to.regress = "percent.mt", verbose = FALSE)
```

```{r}
pbmc <- RunPCA(pbmc, verbose = FALSE)
pbmc <- RunUMAP(pbmc, dims = 1:30, verbose = FALSE)

pbmc <- FindNeighbors(pbmc, dims = 1:30, verbose = FALSE)
pbmc <- FindClusters(pbmc, verbose = FALSE)
DimPlot(pbmc, label = TRUE)
```



# SCTransform Normalization with DoubletFinder
```{r}

```

