---
title: "scRNA w/ Seurat"
output: html_document
date: "2024-02-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library, include=FALSE}
library(dplyr)
library(Seurat)
library(patchwork)
library(sctransform)
library(DoubletFinder)
library(glmGamPoi)
library(presto)
path = "/home/me/r2w2/"
```

## Loading Data

10k Human PBMC, 3' Human peripheral blood mononuclear cells of a healthy female donor aged 25-30.

Libraries were generated from ~16,000 cells (11,984 recovered). Sequenced on an illumina novaseq 6000 to read depth of approximately 40,000 mean reads per cell. 



```{r data_loading0}
# Load the PBMC dataset
pbmc.data <- Read10X(data.dir = paste0(path,"filtered_feature_bc_matrix/"))
# Initialize the Seurat object with the raw (non-normalized data).
pbmc <- CreateSeuratObject(counts = pbmc.data,project="blood_cells",min.cells = 3, min.features = 200)
# store mitochondrial percentage in object meta data
pbmc <- PercentageFeatureSet(pbmc, pattern = "^MT-", col.name = "percent.mt")
```
# Plan: Compare four different pipelines:

For this markdown file we are going to run four pipelines:

I) Seurat Standard
II) Seurat Standard w/ Doublet Finder
III) SCTransform
IV) SCTransform w/ Doublet Finder

# Seurat Standard Pipeline
### Visualizing QC metrics 
### Filtering based on QC metrics
### Normalization
### Feature selection
### Scaling data
### PCA
### KNN
### Clusters
### UMAP/tSNE

```{r violin}
# Visualize QC metrics as a violin plot
VlnPlot(pbmc, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```

```{r scatterplot}
plot1 <- FeatureScatter(pbmc, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(pbmc, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1 + plot2
```

```{r std_seurat}
pbmc <- subset(pbmc, subset = nFeature_RNA > 200 & nFeature_RNA < 80000 & percent.mt < 25)
pbmc <- NormalizeData(pbmc)
pbmc <- FindVariableFeatures(pbmc, selection.method = "vst", nfeatures = 2000)
pbmc <- ScaleData(pbmc)
pbmc <- RunPCA(pbmc)
```


```{r eblow1}
ElbowPlot(pbmc)
```

I would choose a PC of 10 here. 
```{r pca_choice1}
PC_count = 10
pbmc <- FindNeighbors(pbmc,dims=1:PC_count)
pbmc <- FindClusters(pbmc,resolution=1.2)
```
```{r umap1}
pbmc <- RunUMAP(pbmc,dims=1:10)
DimPlot(pbmc,reduction="umap")
```




### DoubletFinder

```{r pK1}
sweep.res.list_blood <- paramSweep(pbmc)
sweep.stats_blood <- summarizeSweep(sweep.res.list_blood)
bcmvn_blood <- find.pK(sweep.stats_blood)

homotypic.prop <- modelHomotypic(pbmc@meta.data$seurat_clusters)
nExp_poi <- round(0.075*nrow(pbmc@meta.data))  ## Assuming 7.5% doublet formation rate - tailor for your dataset
nExp_poi.adj <- round(nExp_poi*(1-homotypic.prop))

pbmc <- doubletFinder(pbmc, PCs = 1:PC_count, pN = 0.25, pK = 0.09, nExp = nExp_poi.adj, reuse.pANN = FALSE, sct = FALSE)
paste0("Doublets found: ",sum(pbmc$DF.classifications_0.25_0.09_847 == "Doublet"))
```
### Markers + Heatmap

```{r markers1}
pbmc.markers <- FindAllMarkers(pbmc, only.pos = TRUE)
pbmc.markers %>%
    group_by(cluster) %>%
    dplyr::filter(avg_log2FC > 1)
```

```{r markers_map1}
pbmc.markers %>%
    group_by(cluster) %>%
    dplyr::filter(avg_log2FC > 1) %>%
    slice_head(n = 10) %>%
    ungroup() -> top10
DoHeatmap(pbmc, features = top10$gene) + NoLegend()
```

### Seurat Standard: Filtered out doublets

```{r sct_doublet}
pbmc <- doubletFinder(pbmc, PCs = 1:PC_count, pN = 0.25, pK = 0.09, nExp = nExp_poi.adj, reuse.pANN = FALSE, sct = FALSE)
paste0("Doublets found: ",sum(pbmc$DF.classifications_0.25_0.09_847 == "Doublet"))

pbmc <- subset(pbmc, subset=DF.classifications_0.25_0.09_847 == "Singlet")
print(length(pbmc$DF.classifications_0.25_0.09_847))
```

```{ markers_Seurat_filtered}
pbmc.markers <- FindAllMarkers(pbmc, only.pos = TRUE)
pbmc.markers %>%
    group_by(cluster) %>%
    dplyr::filter(avg_log2FC > 1)
```

```{r markers_Seurat_filtered_map}
pbmc.markers %>%
    group_by(cluster) %>%
    dplyr::filter(avg_log2FC > 1) %>%
    slice_head(n = 10) %>%
    ungroup() -> top10
DoHeatmap(pbmc, features = top10$gene) + NoLegend()
```



# SCTransform
```{r data_loading1}
# Load the PBMC dataset
pbmc.data <- Read10X(data.dir = paste0(path,"filtered_feature_bc_matrix/"))
# Initialize the Seurat object with the raw (non-normalized data).
pbmc <- CreateSeuratObject(counts = pbmc.data,project="SCT_blood",min.cells = 3, min.features = 200)
# store mitochondrial percentage in object meta data
pbmc <- PercentageFeatureSet(pbmc, pattern = "^MT-", col.name = "percent.mt")
```

### SCT transform + PCA

```{r pca2}
pbmc <- SCTransform(pbmc, vars.to.regress = "percent.mt", verbose = FALSE)

pbmc <- RunPCA(pbmc, verbose = FALSE)
ElbowPlot(pbmc)
```

### Clustering + UMAP

```{r sct_cluster}
PC_count = 13
pbmc <- FindNeighbors(pbmc,dims=1:PC_count)
pbmc <- FindClusters(pbmc,resolution=1.2)

pbmc <- RunUMAP(pbmc, dims = 1:PC_count, verbose = FALSE)
DimPlot(pbmc, label = TRUE)
```



### DoubletFinder ~ SCTransform

```{r pK2}
sweep.res.list_blood <- paramSweep(pbmc)
sweep.stats_blood <- summarizeSweep(sweep.res.list_blood)
bcmvn_blood <- find.pK(sweep.stats_blood)

homotypic.prop <- modelHomotypic(pbmc@meta.data$seurat_clusters)
nExp_poi <- round(0.075*nrow(pbmc@meta.data))  ## Assuming 7.5% doublet formation rate - tailor for your dataset
nExp_poi.adj <- round(nExp_poi*(1-homotypic.prop))

pbmc <- doubletFinder(pbmc, PCs = 1:PC_count, pN = 0.25, pK = 0.09, nExp = nExp_poi.adj, reuse.pANN = FALSE, sct = FALSE)
paste0("Doublets found: ",sum(pbmc$DF.classifications_0.25_0.09_847 == "Doublet"))
```


### Markers

```{r}
pbmc.markers <- FindAllMarkers(pbmc, only.pos = TRUE)
pbmc.markers %>%
    group_by(cluster) %>%
    dplyr::filter(avg_log2FC > 1)
```

```{r}
pbmc.markers %>%
    group_by(cluster) %>%
    dplyr::filter(avg_log2FC > 1) %>%
    slice_head(n = 10) %>%
    ungroup() -> top10
DoHeatmap(pbmc, features = top10$gene) + NoLegend()
```

### SCTransform + Filtered Doublets


```{r filtering_doublets}
pbmc <- subset(pbmc, subset=DF.classifications_0.25_0.09_847 == "Singlet")
print(length(pbmc$DF.classifications_0.25_0.09_847))
```


```{r}
pbmc.markers <- FindAllMarkers(pbmc, only.pos = TRUE)
pbmc.markers %>%
    group_by(cluster) %>%
    dplyr::filter(avg_log2FC > 1)
```

```{r}
pbmc.markers %>%
    group_by(cluster) %>%
    dplyr::filter(avg_log2FC > 1) %>%
    slice_head(n = 10) %>%
    ungroup() -> top10
DoHeatmap(pbmc, features = top10$gene) + NoLegend()
```







